---
title: "citibike_predictions.sh"
output: html_document
date: "2023-06-19"
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Setup


```{r setup}
library(ISLR2)

library(tidyverse)
library(scales)

library(modelr)
options(na.action = na.warn)

library(broom)
library(here)

set.seed(1)
```

## Loading Data

```{r load data}
holiday_data <- read_csv(here("week4/holidays.csv")) %>%
  mutate(date=date(date))
original_data <- read.delim(here("week4/trips_per_day.tsv"), sep ="\t", header=TRUE)
citibike_data <- original_data %>%
  mutate(date=ymd(ymd)) %>% 
  left_join(holiday_data) %>% 
  replace(is.na(.), FALSE) %>% 
  mutate(holiday=as.logical(holiday)) %>% 
  replace(is.na(.), TRUE)
test_sample <- sample(nrow(citibike_data), nrow(citibike_data) * 0.1)

test_data <- citibike_data[test_sample,]
train_data <- citibike_data[-test_sample,]
```

## First Model(s)

```{r First Models}

# Fit n models: 1 more each degree in 1:n
n <- 20
train_err <- c()
validate_err <- c()
for (i in 1:n) {
  model <- lm(num_trips ~ poly(tmin, i), data=train_data)
  
  train_err[i] <- sqrt(mean((predict(model, train_data) - train_data$num_trips)^2))
}


plot_data <- data.frame(power=1:n, train_err)

plot_data %>% 
  ggplot(aes(x=power, y=train_err)) +
  geom_line()
  

```

## Feature Analysis

### Thoughts

Potential useful features

- Weekend vs Weekday
  - More people on the weekend than weekday?
- Holiday
  - If the day is a holiday then this might mean more people are off -> potential negative effect?
- Temperature obviously
  - Maybe take he average temp for that day (tmin-tmax)/2
- Snow day or not
- Raining or not

#### Analysis of snwd, snow, and prcp

```{r Feature Analysis}
train_data %>% 
  group_by(snow) %>% 
  summarize(trips=sum(num_trips)) %>% 
  ggplot(aes(x=snow, y=trips)) +
  geom_line() +
  ggtitle("Snow Effect")

train_data %>% 
  group_by(snwd) %>% 
  summarize(trips=sum(num_trips)) %>% 
  ggplot(aes(x=snwd, y=trips)) +
  geom_line() +
  ggtitle("snwd Effect")

train_data %>% 
  group_by(prcp) %>% 
  summarize(trips=sum(num_trips)) %>% 
  ggplot(aes(x=prcp, y=trips)) +
  geom_line() +
  ggtitle("prcp Effect")


```
snow and snwd appear to have a huge effect when it is 0, but not really much change greater than 0. So I think having a variable with no_prcp, no_snow, no_snwd will have a great effect.

```{r testing snow, prcp, snwd}
train_data %>% 
  mutate(no_snow=prcp==0, no_snwd=snwd==0, no_prcp=prcp==0) %>% 
  ggplot() +
  geom_bar(aes(x=no_prcp))

train_data %>% 
  ggplot(aes(x=tmin, y=num_trips)) +
  geom_point()
```

#### Analysis on day of the week and holiday

How does weekday affect number of rides that day ?

```{r day analysis}

train_data_with_wday <- train_data %>% 
  mutate(date=ymd(date), wday = wday(date)) %>% 
  mutate(is_weekend=wday %in% c(1,7))
train_data_with_wday %>% 
  ggplot(aes(x=date, y=num_trips, color=as.factor(wday))) +
  geom_smooth(se=F)
train_data_with_wday %>% 
  ggplot(aes(x=date, y=num_trips, color=is_weekend)) +
  geom_point()

train_data_with_wday %>% 
  group_by(holiday) %>% 
  summarize(total_trips=sum(num_trips)) %>% 
  ggplot(aes(x=holiday, y=total_trips)) +
  geom_col()

train_data_with_wday %>% 
  ggplot(aes(x=date, y=num_trips, color=holiday)) +
  geom_point()
```
Holiday doesn't really change the number of trips in a day from the general pattern.
However weekend has a noticeable change in the number of trips people take. 

## Model Testing

Here I will test numerous models with different feature combinations. Features I will try are:

##### Features

- Date (encoded as number)
- is_weekend
- t_average (tmax-tmin)/2
- no_weather (if raining/snowing)

```{r Model Testing}

kfoldTest <- function(model, data, k) {
  k_fold_data <- data %>% 
    mutate(fold = row_number() %% k + 1)
  validate_err = rep(0, k)
  for (i in 1:k) {
    held_out_data <-  k_fold_data %>% 
      filter(fold == i)
    training_data <- k_fold_data %>% 
      filter(fold != i)
    
    model <- lm(model$terms, data=training_data)
    validate_err[i] <- sqrt(mean((predict(model, held_out_data) - held_out_data$num_trips)^2))
  }
  
  data.frame(avg_validation_err=mean(validate_err), "r^2"=1-mean(validate_err)^2/var(held_out_data$num_trips))
}

# Getting the dataframe ready

train_data_all_features <- train_data_with_wday %>% 
  mutate(no_weather=prcp==0 & snwd==0 & prcp==0, date_number=yday(date), t_average=(tmin+tmax)/2)

# Train all features

lm_all_features <- lm(num_trips ~ poly(date_number, 2) + no_weather + t_average + is_weekend, data = train_data_all_features)
kfoldTest(lm_all_features, train_data_all_features, 5)

# Try squaring t_average

lm_all_features_2 <- lm(num_trips ~ poly(date_number, 2) + no_weather + poly(t_average, 2) + is_weekend, data = train_data_all_features)
kfoldTest(lm_all_features_2, train_data_all_features, 5)

lm_all_features_3 <- lm(num_trips ~ poly(date_number, 4) + no_weather + poly(t_average, 4) + is_weekend, data=train_data_all_features)
kfoldTest(lm_all_features_3, train_data_all_features, 5)

train_data_all_features %>% 
  add_predictions(lm_all_features_3) %>% 
  ggplot(aes(x=t_average)) +
  geom_point(aes(y=num_trips)) +
  geom_line(aes(y=pred, color=is_weekend, color=no_weather))

```
It looks like lm_all_features_3 has the best r^2 at ~85%. Here I'll modify it to use the existing data.

```{r Package lm_all_features_3}

final_model <- lm(num_trips ~ poly(I(yday(ymd(ymd))), 4) + I(prcp == 0 & snow == 0 & snwd == 0) + poly(I((tmax+tmin)/2) ,4) + I(wday(ymd(ymd)) %in% c(1,7)), data=original_data)

kfoldTest(final_model, original_data, 5)

# Run on test data

1- mean((predict(final_model, test_data) - test_data$num_trips)^2)/var(test_data$num_trips)

sqrt(mean((predict(final_model, test_data) - test_data$num_trips)^2))
identity <- function(X) {x}
original_data %>% 
  add_predictions(final_model) %>% 
  ggplot(aes(x=pred, y=num_trips)) +
  geom_point() +
  stat_function(fun=identity)
save(final_model, file = "model.RData")
```

With my final model I acheived a rmse of 4459 and ~85% R^2 on the test data.

