---
title: "RegressionPractice.rmd"
output: html_document
date: "2023-06-13"
---
# Setup
```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)

library(scales)
library(broom)
library(modelr)
library(tidyverse)

options(na.action = na.warn)

theme_set(theme_bw())
options(repr.plot.width=4, repr.plot.height=3)
```

# ISRS Problem 5.29
```{r Problem 5.29}
body_data <- read.table("body.dat.txt", header = FALSE) %>%
  rename(weight=V23, height=V24)

model = lm(weight ~ height, body_data)
tidy(model)
glance(model)

# This creates the model predictions for each unique point in the data set (independents) (This isn't used after this)
model_predictions <- body_data %>%
  data_grid(height) %>%
  add_predictions(model)

# This creates the plot data
plot_data <- body_data %>%
  group_by(height) %>%
  summarize(count=n(), mean_weight=mean(weight)) %>%
  add_predictions(model)

plot_data %>%
  ggplot(aes(x=height, y=mean_weight)) +
  geom_point(aes(size=count)) +
  geom_line(aes(y=pred))

```

#ISRS Problems 6.1, 6.2, 6.3

```{r Problems 6.1, 6.2, 6.3}
baby_weights <- read.table("babyweights.txt", header=TRUE)
baby_weight_model_smoke <- lm(bwt ~ smoke, baby_weights)
tidy(baby_weight_model_smoke)

baby_weight_model_parity <- lm(bwt ~ parity, baby_weights)
tidy(baby_weight_model_parity)

baby_weight_model_all <- lm(bwt ~ ., baby_weights)
tidy(baby_weight_model_all)


```
6.1

The equation of the regression line is baby_weight = 123.05 - 8.94*smoke

If the mother is a smoker then the baby's weight will decrease by 8.94. 
so the predicted weight of a baby born to a smoker mother is 123.05 - 8.94 = 114.11
otherwise its 123.05.
Yes there is a statistically significant relationship, since the p-value is basically 0.


6.3

b) For every one unit of gestation, bwt increases by .44. For every one year, bwt decreases by 0.01. 

c) Why would there be a difference in the coefficients of the same variables when we include different variables in 
the linear model ?

Because the variables are related to each other? So a higher value variable could cause (maybe not the right word) 

